
#Importing necessary libraries
```{r}
library(urca)
library(tseries)
library(astsa)
library(tsDyn)
```

#Checking the class of chosen data
```{r}
data("econ5")
unemp<-econ5[,1]
gnp<-econ5[,2]
class(unemp)
class(gnp)
ts.plot(unemp)
ts.plot(gnp)
```
#Checking for stationarity of the two series
#H0:Data is non-stationary
#H1:Data is stationary
```{r}
adf.test(unemp)
adf.test(gnp)
```
#p-value for both >0.05, we fail to reject H0.

#Combining the two series
```{r}
data<-cbind(unemp,gnp)
```

#Selecting the appropriate lag/ order of integration
```{r}
lags.select(data)
```

#JOHANSEN COINTEGRATION TEST
#H0:no cointegration
#H1:presence of cointegration
#Type trace
```{r}
test1<-ca.jo(data,type="trace",K=3)
summary(test1)
```
#for r <= 1:
#Null hypothesis: The rank of the cointegration matrix is less than or equal to 1, indicating the presence of at most one cointegration relationship between the variables.
#Alternative hypothesis: The rank of the cointegration matrix is greater than 1, indicating the presence of more than one cointegration relationship between the variables.
#for r = 0:
#Null hypothesis: The rank of the cointegration matrix is 0, indicating no cointegration relationship between the variables.
#Alternative hypothesis: The rank of the cointegration matrix is greater than 0, indicating the presence of at least one cointegration relationship between the variables.
#At 5% level of significance- at most one cointegration relationship between the variables

#Type eigen
```{r}
test2<-ca.jo(data,type="eigen",K=3)
summary(test2)
```
#At 5% level of significance- there exists at most one cointegration relationship among the variables

