---
title: "CAT2_2248111"
author: "Nishanth Girish"
date: "2023-04-27"
output: word_document
---

```{r}
#Importing the necessary libraries:
library(tseries)
library(astsa)
library(forecast)
library(seastests)
```

```{r}
#Reading the data:
data <- read.csv(file.choose())

#Extracting the target column and converting to a time series object:
data1 <- ts(data$Annual.common.stock.price..U.S...1871.to.1970, start = 1871 , end = 1969, frequency = 1)
data1
```


```{r}
#Time series and ACF plot:
ts.plot(data1 , ylab="Stock price", main ="Time series plot")

Acf(data1, main =" ACF plot")
```

INTERPRETATION:

From the time series plot we can observe that there is a trend component. A seasonal component is not present. The model is additvie in nature [Zt = mt + et]. 
The ACF plot indicates that the ACF does not rapidly decrease to zero as lag increases and some of the ACF goes beyond the threshold limit. 
The data is not stationary.




```{r}
#Method of moving average of order 3:
ma1 <- ma(data1, order =3)


error <- data1- ma1
ts.plot(error,ylab="Stock prices",  main = "Time series plot after MA smoothing")

```
INTERPRETATION:
We have applied the method of moving average smoothing of order 3 and removed the trend component from the data as evident fromt the time series plot.

```{r}
#length of the error object:
length(error)
```

```{r}
#Removing the missing values:
newerror <- error[-1]
newerror1 <- newerror[-98]

#ACF and PACF plots after MA smoothing
acf(newerror1, main = "ACF plot after MA smoothing") #tells us about the order of MA. MA(1)
pacf(newerror1,main = "PACF plot after MA smoothing") #tells us about the order of AR. AR(3)

```

INTERPRETATION:

The ACF plot indicates that 2 lags are significant and the PACF plot indicates that 3 lags are significant.
From the ACF and PACF plots of the detrended data, we suspect an ARMA(3,2) process.


```{r}

# Automatically fitting an  ARMA model using in-built function:
fit <- auto.arima(newerror1,seasonal=FALSE)     
summary(fit)
```

INTERPRETATION:

Using the in-built function in R, we get an ARMA(0,2) process with moving average process parameters theta1 = -0.7464  and theta2 = -0.1924.
AIC value of the model is 313.06 and BIC of the model is 320.78. The variance of the model is 1.388.
RMSE is 1.1659



```{r}
# We suspect an ARMA of different orders from the PACF and ACF plots.

# Fitting ARMA models based on the ACF and PACF plots of the stationary data:
fit2 <- arima(newerror1, order = c(3,0,2))
summary(fit2)
```

INTERPRETATION:
We fit ARMA(3,2) model with AR parameters phi1, phi2, phi3:  -0.7052, 0.0228, -0.2378 respectively,
and MA parameters theta1 and theta2: -0.1013,  -0.8986 respectively.
The AIC value of the model is 313.42 and the variance of the model is 1.215. RMSE is 1.1023.

```{r}
# We suspect an ARMA of different orders from the PACF and ACF plots.

# Fitting ARMA models based on the ACF and PACF plots of the stationary data:
fit3 <- arima(newerror1, order = c(3,0,1))
summary(fit3)
```

```{r}
#Residual analysis for model 1:
r = resid(fit)

acf(r, main = "ACF plot of the residuals")            #non-significant acf at different lags.

#Shapiro-Wilk test to check the normality of the residuals:
#H0: Residuals are normal.
#H1: Residuals are not normal.
shapiro.test(r)  #p-value < 0.05 -> null hypothesis 
#Q-Q plot:
qqnorm(r)
qqline(r)

#Box-Pierce test to examine the null hypothesis of independence.
#H0: There is no significant autocorrelation.
#H1: There is some significant autocorrelation.
Box.test(r,lag = 2) #lag = p + q; p=0, q= 2

```

INTERPRETATION:

1. From the Shapiro-Wilk test, we get p-value < 0.05. Hence, we reject the null hypothesis that the errors are normally distributed and conclude that the errors are not normally distributed.

2. The Q-Q plot also depicts that some of the errors are far from the reference line and hence we conclude that the errors are not normal.

3. From the Box-Pierce test, we get p-value 0.2781. Hence, we fail to reject the null hypothesis that there is no significant autocorrelation.

4. ACF plot indicates that there are no significant ACF values. Errors are uncorrelated.

```{r}

#Residual analysis for model 2:
r = resid(fit2)

acf(r, main = "ACF plot of the residuals")            #non-significant acf at different lags.

#Shapiro-Wilk test to check the normality of the residuals:
#H0: Residuals are normal.
#H1: Residuals are not normal.
shapiro.test(r)  #p-value < 0.05 -> null hypothesis 
#Q-Q plot:
qqnorm(r)
qqline(r)

#Box-Pierce test to examine the null hypothesis of independence.
#H0: There is no significant autocorrelation.
#H1: There is some significant autocorrelation.
Box.test(r,lag = 5) #lag = p + q; p=3, q= 2


```

INTERPRETATION:

1. From the Shapiro-Wilk test, we get p-value < 0.05. Hence, we reject the null hypothesis that the errors are normally distributed and conclude that the errors are not normally distributed.

2. The Q-Q plot also depicts that some of the errors are far from the reference line and hence we conclude that the errors are not normal.

3. From the Box-Pierce test, we get p-value 0.9111. Hence, we fail to reject the null hypothesis that there is no significant autocorrelation.

4. ACF plot indicates that there are no significant ACF values. Errors are uncorrelated.

CONCLUSION:
The second model performs slightly better than the first model as variance and RMSE value is less. There is not much difference between the AIC values of both the models.