---
title: "MST371_CAT1"
author: "Nishanth Girish"
date: '2023-04-03'
output: word_document
---

```{r}
#Importing necessary libraries:
library(forecast)
library(seastests)
library(astsa)
library(tseries)
```

DATASET DESCRIPTION:
The given dataset is the 'globtemp' dataset which is contained in the astsa library of R software. The data contained is the global mean land-ocean temperature deviations (from 1951-1980 average), measured in degrees centigrade, for the years 1880-2015.

```{r}
#Loading the dataset:
data <- globtemp
data
```


```{r}
#Time series plot:
ts.plot(data, ylab ='Temperature deviation', main = 'Time series plot')

#ACF Plot:
acf(data, main = 'ACF Plot')
```

INTERPRETATION:

From the time series plot, it is clear that the data follows an increasing trend. No seasonal component can be observed. Hence, it is an additive model [Zt = mt + et].
From the ACF plot, it can be observed that the ACF is not rapidly decreasing to zero. This implies that the data is not stationary.

```{r}
#Differencing to remove trend component
td <- diff(data)

#Time series and ACF plot afte differencing:
ts.plot(td, ylab ='Temperature deviation', main = 'Time series plot after differencing:')
acf(td, main = 'ACF plot after differencing:')
```
INTERPRETATION:

The method of differencing was applied to remove the trend component from the data and make it stationary.
The time series plot after differencing does not show trend or seasonal variation and appears to be random in nature.
The ACF plot after differencing shows that the ACF rapidly drops to zero after a few lags indicating that there is no trend or seasonal component. It can be concluded that the data is now stationary. 

```{r}
#PACF Plot:
pacf(td , main =' PACF Plot:')
```

INTERPRETATION:

The PACF plot shows that the PACF cuts off after 3 lags. Thus, AR(3) process can be suspected.

```{r}
#Give whatever order that we suspect from the PACF

#Fitting an AR(3) model
fit <- arima(td,order = c(3,0,0))
fit
```

INTERPRETATION:

An AR(3) model was fitted onto the data. 
The autoregressive parameters obtained are: -0.3669,  -0.3421,  -0.2363.
The intercept is 0.0071.


