```{r}
library(tseries)
library(forecast)
library(astsa)
library(vars)
```
#About the Dataset(CANADA)

This dataset contains information about the Canadian Labour Productivity(prod), Employment Rate(e), Unemployment Rate(u) and the Real Wages(rw) from the year 1980 to 2000 and is a Quarterly Data.

```{r}
#Downloading the data
data=Canada
head(data)
```
```{r}
#Checking if Time Seires Data or not
class(data)

#We see that it is a Multivariate Time Series.
```

```{r}
#TIME SERIES PLOT
par(mfrow=c(2,2))

ts.plot(data[,1],ylab= 'Employment', main="Employment")
ts.plot(data[,2],ylab= 'Productivity', main="Productivity")
ts.plot(data[,3],ylab= 'Real Waget', main="Real Wage")
ts.plot(data[,4],ylab= 'Unmployment', main="Unemployment")

#We see that trend exists in some series of our data, we check stationarity using the ACF PACF PLOTS.
```

```{r}
#ACF/PACF of all individual series 

par(mfrow=c(2,2))
for(i in 1:ncol(data)){
  Acf(data[,i], main=paste("ACF -", names(data)[i]))
  pacf(data[,i], main=paste("PACF -", names(data)[i]))
}

#Conclusion: We observe that the ACF plots of all the 4 variables are not rapidly decreasing to 0, hence they are not stationary.
```

```{r}
#Converting to Stationary using DIFFERENCING
data1=diff(data)

#plotting tsplot 
par(mfrow=c(2,2))

ts.plot(data1[,1],ylab= 'Employment',main="Employment")
ts.plot(data1[,2],ylab= 'Productivity',main="Productivity")
ts.plot(data1[,3],ylab= 'Real Wage',main="Real Wage")
ts.plot(data1[,4], ylab= 'Unemployment',main="Unemployment")

#We observe randomness, we confirm stationarity using ACF/PACF plots
```

```{r}
#ACF/PACF of all individual series after differencing

par(mfrow=c(2,2))
for(i in 1:ncol(data1)){
  Acf(data1[,i], main=paste("ACF -", names(data1)[i]))
  pacf(data1[,i], main=paste("PACF -", names(data1)[i]))
}

#Conclusion: We observe that the ACF plots of all the 4 variables are rapidly decreasing to 0, hence they are stationary.
```
#VAR(P) MODELING

Now in a VAR(p) model we need to know how many lags(p) to be considered for the modeling.
Through the ACF/PACF plots we see that most of the variable have a significant correlation upto lag 2. Hence we can consider p=2 for our modelling.

But to check whether this is the best model ,we fit the VAR(p) model for p=1,2,3 and then check the AIC values to find the best model.

```{r}
#Fitting the model for p=1,2,3
VAR1=VAR(data1,p=1)
VAR2=VAR(data1,p=2)
VAR3=VAR(data1,p=3)

#Checking AIC values
AIC(VAR1)
AIC(VAR2)
AIC(VAR3)

#Conclusion: We see that the AIC value for VAR(2) model is the least hence we move forward with it.
```

```{r}
#Getting the Coefficients

VAR2

#Interpretation: For each of the 4 variables in our multivariate time series, we get the estimated value of coefficients. Since lag=2 was considered in our model,each variable is based on the previous 2 lags of all the 4 variables.
```
```{r}
#FORECASTING

predict(VAR2,n.ahead = 4)

#Now these predicted values are the Differenced predicted values. We need to transform this back to get the actual differenced values.
```
#############################################################################



